---
title: "Geo Lines Plot"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- ANY PACKAGES USED HERE MUST BE LOADED IN THE DESCRIPTION FILE AS WELL !-->

# Overview of function

`geo_marker_plot` can be used to create the following kinds of charts:

- Geo marker plot

Geo marker plots contain markers that highlight important geographic positions, the markers may differ in colour but do not use size as a channel for communicating data. For charts where geographic positions should be represented as circles of differing sizes, refer to `geo_bubble_chart` instead.

# data_geo_lines_plot

```{r}
library(oidnChaRts)
```

This vignette covers the use of geo lines plots for visualising data with a variety of htmlwidget libraries, for demonstration purposes we use the following dataset generated from https://doi.org/10.6084/m9.figshare.4516772. 

```{r}
head(data_geo_lines_plot)
```

# Geo lines plot specification

Geo lines plot is used to visualise journies between geolocations, specifically a directed relationship between point A and B (i.e. there is no support for multi-legged journies). As all geo lines are directional, the following communication channels are available within the plot:

- start: geo locations from which journies start.
- end: geo locations from which journies end.
- both: geo locations which are both start and end points.

The `geo_lines_plot` determines which locations are both "send" and "receive" locations and labels them appropriately.

# Geo lines plot

The following will create a generic geo bubble chart with leaflet, note that significant work is required to make legends in leaflet charts "active" and you are recommended to build such charts from scratch.

```{r}
library(leaflet)
data_geo_lines_plot %>%
  geo_lines_plot(library = "leaflet")
```

## Popup on geolines

```{r}
library(leaflet)
data_geo_lines_plot %>%
  geo_lines_plot(library = "leaflet",line.popup = ~start.location)
```


## Popup on termini ()

```{r}
library(leaflet)
data_geo_lines_plot %>%
  geo_lines_plot(library = "leaflet",line.popup = ~start.location)
```


