---
title: "Geo Marker Plots"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- ANY PACKAGES USED HERE MUST BE LOADED IN THE DESCRIPTION FILE AS WELL !-->

# Overview of function

`geo_marker_plot` can be used to create the following kinds of charts:

- Geo marker plot

Geo marker plots contain markers that highlight important geographic positions, the markers may differ in colour but do not use size as a channel for communicating data. For charts where geographic positions should be represented as circles of differing sizes, refer to `geo_bubble_chart` instead.

```{r, echo=FALSE}
data_geo_marker_plot %>%
  geo_marker_plot(library = "leaflet",
                  mouseover.text = ~city,
                  popup.text = ~paste("Country:",country, "<br>", "City:", city))
```


As with all `oidnChaRts` libraries, you are advised to load the htmlwidget library you're using directly.

# data_stacked_bar_chart

```{r}
library(oidnChaRts)
```

This vignette covers the use of geo marker plots for visualising data with a variety of htmlwidget libraries, for demonstration purposes we use the following dataset generated from https://doi.org/10.6084/m9.figshare.4516772. 

```{r}
head(data_geo_marker_plot)
```


## Play

```{r}
data_
```





