---
title: "Geo Bubble Chart"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- ANY PACKAGES USED HERE MUST BE LOADED IN THE DESCRIPTION FILE AS WELL !-->

# Overview of function

`geo_bubble_chart` can be used to create the following kinds of charts:

- Geo bubble chart

Geo bubble charts contain circle that highlight important geographic positions, and the radius of this circle markers communicates additional information about these locations. Note that these visualisations can be become difficult to interpet where there are high point densities. For charts where you only wish to communicate the geographic locations of certain points, refer to `geo_marker_plot` instead.

```{r, echo=FALSE}

```


As with all `oidnChaRts` libraries, you are advised to load the htmlwidget library you're using directly.

# data_stacked_bar_chart

```{r}
library(oidnChaRts)
```

This vignette covers the use of geo bubble charts for visualising data with a variety of htmlwidget libraries, for demonstration purposes we use the following dataset generated from https://doi.org/10.6084/m9.figshare.4516772. 

```{r}
head(data_geo_bubble_chart)
```

The data is generated from a historical analysis of the letters sent between German migrants and their families between the 18th and 19th Centuries, the columns may be summarised as follows:

- location.name: Country, City (state, if relevant)
- latitude: Latitude of location
- longitude: Longitude of location
- count: number of times location 

# Geo bubble chart specifications

Geo bubble charts provide multiple channels of communication:

- position: the longitude, latitude of the data point
- radius: size of the bubble
- colour: allows bubbles to be grouped together by kind

There is significant variation in the counts column of the sample dataset:

```{r}
library(tidyverse)
data_geo_bubble_chart %>%
  ggplot(aes(x = "", count)) + geom_violin() + coord_flip()
```

To improve the appearance of the chart, we will use the tidyverse to filter out points below the 50th quartile and use the `scales` library to rescale the counts from 3 to 20:

```{r}
library(scales)
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  ggplot(aes(x = "", count)) + geom_violin() + coord_flip()
```

Additionally, colour groupings will be added to the dataset as follows:

```{r}
data_geo_bubble_chart %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Paired")
  ))
```


# Geo bubble chart

The following will create a generic geo bubble chart with leaflet, note that significant work is required to make legends in leaflet charts "active" and you are recommended to build such charts from scratch.

```{r}
library(leaflet)
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Dark2")
  )) %>%
  geo_bubble_chart(
    library = "leaflet",
    bubble.radius = ~count,
    bubble.group = ~country,
    color = ~color
  )
```

The following will create a generic highchart geo bubble chart, note the addition of `hc_mapNavigation` to enable zooming.

```{r}
library(highcharter)
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Dark2")
  )) %>%
  geo_bubble_chart(
    library = "highcharter",
    bubble.radius = ~count,
    bubble.group = ~country
  ) %>%
    hc_mapNavigation(enabled = TRUE)
```

This will create a generic geo bubble chart using plotly, note that the legend items automatically scale based on the largest size bubble in the series.

```{r}
library(plotly)
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Dark2")
  )) %>%
  geo_bubble_chart(
    library = "plotly",
    bubble.radius = ~count,
    bubble.group = ~country
  )
```


## Mouseover text

Mouseover text is supported in the leaflet library as follows, note that arbitrary information can easily be included in the text.

```{r}
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Dark2")
  )) %>%
  geo_bubble_chart(
    library = "leaflet",
    bubble.radius = ~count,
    bubble.group = ~country,
    mouseover.text = ~paste("Country:", country)
  )
```

It is easiest to add custom mouseover/tooltip content to `highcharter` charts using the `hc_tooltip` function directly:

```{r}
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Dark2")
  )) %>%
  geo_bubble_chart(
    library = "highcharter",
    bubble.radius = ~count,
    bubble.group = ~country
  ) %>%
    hc_mapNavigation(enabled = TRUE) %>%
  hc_tooltip(
        formatter = JS(
          "function(){
console.log(this)
          return '<b>Country: </b>' + this.series.name + 
                  '<br/>' +
                  '<b>City: </b>' + this.point.city +
                  '<br/>' +
                  '<b>Latitude: </b>' + this.point.lat +
                  '<br/>' +
                  '<b>Longitude: </b>' + this.point.lon;}"
          )
        )
```

Thoroughly customising the content of `plotly` tooltips is difficult, but it is possible to add additional information to mouseovers as with the `mouseover.text` argument:

```{r}
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Dark2")
  )) %>%
  geo_bubble_chart(
    library = "plotly",
    bubble.radius = ~count,
    bubble.group = ~country,
    mouseover.text = ~paste("City:", city, "<br>",
                            "Country:", country)
  )
```

## Popup text

The leaflet library supports "popup text", content shown when the bubble is selected. The other libraries available for `geo_bubble_chart` do not support this without using Shiny or crosstalk.

```{r}
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Dark2")
  )) %>%
  geo_bubble_chart(library = "leaflet",
    bubble.radius = ~count,
    bubble.group = ~country,
                  mouseover.text = ~city,
                  popup.text = ~paste("Country:",country, "<br>", "City:", city))
```

## Alternative base maps (projections, etc)

The leaflet library has a very flexible system for modifying the base maps and projections in visualisations, and this can be done "after the fact", i.e. the `geo_bubble_chart` can be used to generate the template map and other base maps applied.

```{r}
data_geo_bubble_chart %>%
  filter(count > 5) %>%
  mutate(count = rescale(count, to = c(3,20))) %>%
  mutate(color = plyr::mapvalues(
    country,
    from = unique(country),
    to = RColorBrewer::brewer.pal(length(unique(country)), "Dark2")
  )) %>%
  geo_bubble_chart(
    library = "leaflet",
    bubble.radius = ~count,
    bubble.group = ~country
  ) %>%
  addProviderTiles(providers$Esri.NatGeoWorldMap)
```

The `highcharter` library requires the base map and other details to be specified in the original function call, the appriopriate documentation to refer to is here: http://jkunst.com/highcharter/highmaps.html

The `plotly` library also requires base maps and other geographic details to be specified in the original function call.


