---
title: "Geo Marker Plots"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- ANY PACKAGES USED HERE MUST BE LOADED IN THE DESCRIPTION FILE AS WELL !-->

# Overview of function

`geo_marker_plot` can be used to create the following kinds of charts:

- Geo marker plot

Geo marker plots contain markers that highlight important geographic positions, the markers may differ in colour but do not use size as a channel for communicating data. For charts where geographic positions should be represented as circles of differing sizes, refer to `geo_bubble_chart` instead.

```{r, echo=FALSE}
data_geo_marker_plot %>%
  geo_marker_plot(library = "leaflet",
                  mouseover.text = ~city,
                  popup.text = ~paste("Country:",country, "<br>", "City:", city))
```


As with all `oidnChaRts` libraries, you are advised to load the htmlwidget library you're using directly.

# data_stacked_bar_chart

```{r}
library(oidnChaRts)
```

This vignette covers the use of geo marker plots for visualising data with a variety of htmlwidget libraries, for demonstration purposes we use the following dataset generated from https://doi.org/10.6084/m9.figshare.4516772. 

```{r}
head(data_geo_marker_plot)
```

The data is generated from a historical analysis of the letters sent between German migrants and their families between the 18th and 19th Centuries, the columns may be summarised as follows:

- location.name: Country, City (state, if relevant)
- latitude: Latitude of location
- longitude: Longitude of location
- count: number of times location 

# Geo marker plot specifications

In Geo marker plots the only channels of information (i.e. what can vary between data points) are as follows:

- position: the longitude, latitude of the data point
- colour: colour of the markers

However, it must be noted that geo marker icons are typcially implemented as .png images and therefore it is not generally possible to have **n** different colours of markers. For this reason, the `geo_marker_plot` function provides absolutely minimal flexibility in the appearance of your output.

At the time of writing (`r Sys.Date()`) the only library with good support for geo markers is `leaflet`. You're strongly recommended to build these charts from scratch, following the documentation here: [http://rstudio.github.io/leaflet/markers.html](http://rstudio.github.io/leaflet/markers.html).

# Geo marker plot

The following will create a generic geo marker plot with leaflet:

```{r}
data_geo_marker_plot %>%
  geo_marker_plot(library = "leaflet")
```

## Mouseover text

Mouseover text is supported in the leaflet library as follows:

```{r}
data_geo_marker_plot %>%
  geo_marker_plot(library = "leaflet",
                  mouseover.text = ~country)
```

## Popup text

Popup text (which is shown when a marker is clicked) is supported for leaflet as follows:

```{r}
data_geo_marker_plot %>%
  geo_marker_plot(library = "leaflet",
                  mouseover.text = ~city,
                  popup.text = ~paste("Country:",country, "<br>", "City:", city))
```





