---
title: "Charts"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- ANY PACKAGES USED HERE MUST BE LOADED IN THE DESCRIPTION FILE AS WELL !-->
# Overview of function

`line_chart` can be used to create the following kinds of charts:

- Line charts

As with all `oidnChaRts` libraries, you are advised to load the htmlwidget library you're using directly.

# data_stacked_bar_chart

```{r}
library(oidnChaRts)
```

This vignette covers the use of line charts for visualising data with a variety of htmlwidget libraries, for demonstration purposes we use the following dataset generated from https://doi.org/10.6084/m9.figshare.3761562. The dataset concerns the number of jobs advertised on a number of "freelance websites" on a specific date, here's how the data looks:

```{r}
library(oidnChaRts)
```

This is our dataset:

Initial call

```{r}
library(highcharter)
line_chart(data = data_line_chart,
           library = "highcharter",
           x.column = ~af,
           y.column = ~fdr,
           traces.column = ~series) %>%
  hc_yAxis(
        type = "logarithmic",
        title = list(text = "FDR"),
        labels = list(
          formatter = JS(
            "function () {
            var label = this.axis.defaultLabelFormatter.call(this);
            return this.value.toExponential(2);
            }"
))) %>%
  hc_plotOptions(series = list(marker = list(symbol = "circle")))
```

