---
title: "Maps"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Geolines Map

Our sample dataset is included in the library, `data_geo_lines_map` and comprises a set of send-receive coordinates. The `geo_lines_map` function requires a dataset with at least the following columns:

- sender.latitude
- sender.longitude
- receiver.latitude
- receiver.longitude

```{r}
library(oidnChaRts)
head(data_geo_lines_map)
```

Hat tip to [http://personal.tcu.edu/kylewalker/interactive-flow-visualization-in-r.html](http://personal.tcu.edu/kylewalker/interactive-flow-visualization-in-r.html) for the `geospheres` library which computes [great circles](https://en.wikipedia.org/wiki/Great_circle).

## Leaflet

### Basic Example

The minimal number of arguments required for `geo_line_map` are as follows:

```{r}
library(leaflet)
geo_lines_map(data_geo_lines_map, 
              library = "leaflet")
```

There are additional arguments that may control the `line.options` and `termini.options` (the dots at each end of the geo_line)

```{r}
geo_lines_map(data_geo_lines_map, 
              library = "leaflet",
              line.options = list(weight = 3))
```

The leaflet map produced by `geo_lines_map` can be used in pipe chains like any other leaflet map:

```{r}
geo_lines_map(data_geo_lines_map, 
              library = "leaflet",
              line.options = list(weight = 3)) %>%
  addProviderTiles("Stamen.Watercolor")
```

### Tooltips

The `geo_lines_map` function has been designed for tooltips to be easily added to the lines of termini, in the example below a popup will show when a geo_line is clicked.

```{r}
label_journey <- function(sender.location = NA, receiver.location = NA){
  paste0(
    "<p>Sender: ", sender.location,
    "</p>",
    "<p>Receiver: ", receiver.location
  )
}
geo_lines_map(data_geo_lines_map, 
              library = "leaflet",
              line.popup = ~label_journey(sender.location, receiver.location),
              line.options = list(weight = 3)) %>%
  addProviderTiles("Stamen.Watercolor")
```


# Choropleth from points

Our sample dataset is included in the library, `data_geo_points` and comprises a set of coordinates . The `geo_lines_map` function requires a dataset with at least the following columns:

- sender.latitude
- sender.longitude
- receiver.latitude
- receiver.longitude

```{r}
library(oidnChaRts)
head(data_geo_lines_map)
