---
title: "Overview of oidnChaRts"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- ANY PACKAGES USED HERE MUST BE LOADED IN THE DESCRIPTION FILE AS WELL !-->
# What is oidnChaRts useful for?

oidnChaRts is designed to support the following use cases:

- Provide advice on which charts to use for each "kind" of data
- Allow R users to easily compare how different htmlwidgets display different kinds of charts
- Simple tool for creating generic charts

You might find that the output of oidnChaRts is sufficiently useful for your visualisation needs, but it's specifically not designed to create an infinitely flexible manner to create chart X in different libraries - because that would be very difficult.

# Combining oidnChaRts together

All good `htmlwidget` libraries use the `%>%` (pipe) operator to provide a consistent method to construct/add layers to charts. This is achieved by insisting on the first argument of most funtions being a htmlwidget of the kind you're building. 

As an example, consider a `htmlwidget` library called `pokedexeR` which creates a display of your Pokémon data using the following functions:

- pokedex; creates a htmlwidget of the kind `pokedexeR` containing your Pokémon data
- pokedexRegion; specifies the origin of the Pokédex to use, each region has a very distinctive formfactor
- pokedexResolution; specifies the effective resolution of the Pokédex display

These functions would be piped together as follows:

```{r, eval=FALSE}
my_pokemon_data %>%
  pokedex() %>%
  pokedexRegion(region = "Kanto") %>%
  pokedexResolution(resolution = "160x144")
```

The functions in the `oidnChaRts` library that output charts are specifically designed **not** to support this workflow, i.e. one cannot do the following:

```{r, eval = FALSE}
## This will fail, please read the explanation above
my_data %>%
  scatter_plot(library = "highcharter") %>% # function from oidnChaRts
  line_chart(library = "highcharter") %>% # function from oidnChaRts
  hc_tooltip(shared = TRUE) # function directly from highcharter
```

However, in general one can take the output of a `oidnChaRts` function and use it within the htmlwidget library's own functions:

```{r}
library(leaflet)
data_geo_lines_plot %>%
  geo_lines_plot() %>%
  addMarkers(lat = 0, lng = 0, label = "Null Island!")
```








