---
title: "Stacked bar chart"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- ANY PACKAGES USED HERE MUST BE LOADED IN THE DESCRIPTION FILE AS WELL !-->

# Load the library

```{r}
library(oidnChaRts)
```

This vignette covers the following charts:

- Grouped bar charts
- Stacked bar charts
- Scatterplots
- Line charts
- Treemaps

# Grouped Bar Charts

The data for grouped bar charts is the same as that for all bar charts:

```{r}
data_stacked_bar_chart
```

The following will create a generic grouped bar chart with highcharter:

```{r}
stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "highcharter",
                  categories.column = ~country,
                  subcategories.column = ~activity,
                  value.column = ~hours)
```

The following will create a generic grouped bar chart with plotly:

```{r}
stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "plotly",
                  categories.column = ~country,
                  subcategories.column = ~activity,
                  value.column = ~hours)
```

## Ordering Categories

Category order can be supplied directly to the `country_order` column.

```{r}
country_order <- c("United Kingdom", "Canada", "China", "Estonia", "Egypt", "France", "Germany", 
"Ireland", "Saudi Arabia")
stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "highcharter",
                  categories.column = ~country,
                  categories.order = country_order,
                  subcategories.column = ~activity,
                  value.column = ~hours)
```

Plotly ordered categories:

```{r}
country_order <- c("Canada", "China", "Egypt", "Estonia", "France", "Germany", 
"Ireland", "Saudi Arabia", "Slovakia", "United Kingdom")
stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "plotly",
                  categories.column = ~country,
                  categories.order = country_order,
                  subcategories.column = ~activity,
                  value.column = ~hours)
```

## Explicitly Ordering Subcategories

Highcharter explicitly ordered subcategories:

```{r}
activity_order <- c("Unclassified", "Business", "Overlay", "Personal Visit", "Teleconference")

stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "highcharter",
                  categories.column = ~country,
                  categories.order = country_order,
                  subcategories.column = ~activity,
                  value.column = ~hours,
                  subcategories.order = activity_order)
```

Plotly explicitly ordered subcategories:

```{r}
activity_order <- c("Unclassified", "Business", "Overlay", "Personal Visit", "Teleconference")

stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "plotly",
                  categories.column = ~country,
                  categories.order = country_order,
                  subcategories.column = ~activity,
                  value.column = ~hours,
                  subcategories.order = activity_order)
```

# Stacked Bar Charts

The data for grouped bar charts is the same as that for all bar charts:

```{r}
data_stacked_bar_chart
```

If highcharter is selected, the data will automatically be reshaped into wide format:

```{r}
stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "highcharter",
                  categories.column = ~country,
                  subcategories.column = ~activity,
                  value.column = ~hours,
                  stacking.type = "normal")
```

If plotly is selected the data will be provided directly to plotly for visualisation:

```{r}
stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "plotly",
                  categories.column = ~country,
                  subcategories.column = ~activity,
                  value.column = ~hours,
                  stacking.type = "normal")
```



```{r}
activity_order <- c("Unclassified", "Business", "Overlay", "Personal Visit", "Teleconference")

stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "highcharter",
                  categories.column = ~country,
                  categories.order = country_order,
                  subcategories.column = ~activity,
                  value.column = ~hours,
                  subcategories.order = activity_order,
                  stacking.type = "normal")
```

Plotly explicitly ordered subcategories with stacked bars, note that legend order cannot be fixed (https://github.com/plotly/plotly.js/issues/394)

```{r}
activity_order <- c("Unclassified", "Business", "Overlay", "Personal", "Teleconference")

stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "plotly",
                  categories.column = ~country,
                  categories.order = country_order,
                  subcategories.column = ~activity,
                  value.column = ~hours,
                  subcategories.order = activity_order,
                  stacking.type = "normal")
```



## Stacked Percentage

Highcharter explicitly ordered subcategories with stacked bars.

```{r}
activity_order <- c("Unclassified", "Business", "Overlay", "Personal", "Teleconference")

stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "highcharter",
                  categories.column = ~country,
                  categories.order = country_order,
                  subcategories.column = ~activity,
                  value.column = ~hours,
                  subcategories.order = activity_order,
                  stacking.type = "percent")
```

Plotly explicitly ordered subcategories with stacked bars, note that legend order cannot be fixed (https://github.com/plotly/plotly.js/issues/394)

```{r}
activity_order <- c("Unclassified", "Business", "Overlay", "Personal", "Teleconference")

stacked_bar_chart(data = data_stacked_bar_chart,
                  library = "plotly",
                  categories.column = ~country,
                  categories.order = country_order,
                  subcategories.column = ~activity,
                  value.column = ~hours,
                  subcategories.order = activity_order,
                  stacking.type = "percent")
```