---
title: "Stacked bar chart"
author: "Martin John Hadley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- ANY PACKAGES USED HERE MUST BE LOADED IN THE DESCRIPTION FILE AS WELL !-->

# Overview of function

`stacked_bar_chart` can be used to create the following kinds of charts:

- Grouped bar charts
- Stacked bar charts
- Percentile stacked bar charts

As with all `oidnChaRts` libraries, you are advised to load the htmlwidget library you're using directly.

# data_stacked_bar_chart

```{r}
library(oidnChaRts)
```

This vignette covers the use of stacked bar charts for visualising data with a variety of htmlwidget libraries, for demonstration purposes we use the following dataset generated from https://doi.org/10.6084/m9.figshare.3761562. The dataset concerns the number of jobs advertised on a number of "freelance websites" on a specific date, here's how the data looks:

```{r}
head(data_stacked_bar_chart)
```

The original dataset includes measurements since 2016/09/05, this snapshot is re-generated everytime the `oidnChaRts` library is rebuilt, the current value is `r data_stacked_bar_chart$timestamp[1]`. The columns of interest are as follows:

- country: where the employer for a job is located
- country_group: groupings determined by the researchers, i.e. "other Europe" and "all Africa"
- occupation: classification of the job posting
- count: how many jobs met the above specifications

# Stacked barchart specifications

In stacked/grouped bar chart there are "categories" and "subcategories" of observations, here are possible combinations for our dataset:

- Categories: country_group, Subcategories: country, Purpose of chart: Which countries are most significant within each country grouping?
- Categories: country_group, Subcategories: occupations, Question chart answers: Which occupations are most signficiant within each country_group grouping?

In order to create these charts, we must first process the data using the tidyverse:

```{r}
data_stacked_bar_chart %>%
  group_by(country_group, country) %>%
  summarise(total = sum(count)) %>%
  ungroup()
```

This code will be used in all further examples in this document, note that for your analyses it might make more sense to store the output of this code as a new symbol.


# Grouped Bar Charts

Grouped barcharts group data into categories and then display a separate bar for each subcategory.

```{r, echo=F}
descending_order_of_occupations <- data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  mutate(total = sum(count)) %>%
  mutate(total.in.group = sum(total)) %>%
  arrange(desc(total.in.group)) %>%
  ungroup() %>%
  select(occupation) %>%
  unique() %>%
  .[[1]]

descending_order_of_regions <- data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  mutate(total.in.group = sum(total)) %>%
  arrange(desc(total.in.group)) %>%
  select(country_group) %>%
  unique() %>%
  .[[1]]

data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    subcategories.order = descending_order_of_occupations,
                    value.column = ~total)
remove(descending_order_of_occupations, descending_order_of_regions)
```


The following will create a generic grouped bar chart with highcharter:

```{r}
library(highcharter)
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    categories.column = ~country_group,
                    subcategories.column = ~occupation,
                    value.column = ~total)
  
```

The following will create a generic grouped bar chart with plotly:

```{r}
library(plotly)
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "plotly",
                    categories.column = ~country_group,
                    subcategories.column = ~occupation,
                    value.column = ~total)
```

At the time of writing `r Sys.Date()`, the plotly library does a poor job at ensuring the category labels (y-axis labels) are not chopped off. The `layout` function from `plotly` allows us to modify the margins and drop the unnecessary label for the y-axis.

```{r}
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "plotly",
                    categories.column = ~country_group,
                    subcategories.column = ~occupation,
                    value.column = ~total) %>%
  layout(margin = list(l = 150),
         yaxis = list(title = ""))
```


## Ordering Categories

Category order can be supplied directly to the `categories.order` column.

```{r}
## Note: future versions of the dataset may use different groupings, this code assumes "United States" and "United Kingdom" remain distinct entities in "country_group"
order_of_regions <- c(c("United States", "United Kingdom"), setdiff(unique(data_stacked_bar_chart$country_group), c("United States", "United Kingdom")))
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    categories.column = ~country_group,
                    categories.order = order_of_regions,
                    subcategories.column = ~occupation,
                    value.column = ~total)
```

A good pattern to follow with grouped bar charts is to order the categories from "largest to smallest", i.e. the country_group with the most number of jobs at the top of the chart and the least at the bottom. This functionality is deliberately NOT provided by the `stacked_bar_chart` function, instead you *must* modify the data:

```{r}
descending_order_of_regions <- data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  mutate(total.in.group = sum(total)) %>%
  arrange(desc(total.in.group)) %>%
  select(country_group) %>%
  unique() %>%
  .[[1]]

data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "plotly",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    value.column = ~total) %>%
  layout(margin = list(l = 150),
         yaxis = list(title = ""))
```


## Explicitly Ordering Subcategories

Subcategory order can be supplied directly to the `subcategories.order` column.

```{r}
## Note: future versions of the dataset may use different occupations, this code assumes "Writing and translation" and "Creative and multimedia" remain distinct entities in "occupation"
order_of_occupations <- c(c("Writing and translation", "Creative and multimedia"), setdiff(unique(data_stacked_bar_chart$occupation), c("Writing and translation", "Creative and multimedia")))
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    subcategories.order = order_of_occupations,
                    value.column = ~total)
```

In addition to ordering the categories by descending size, it is useful to do the same for the subcategories:

```{r}
descending_order_of_occupations <- data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  mutate(total = sum(count)) %>%
  mutate(total.in.group = sum(total)) %>%
  arrange(desc(total.in.group)) %>%
  ungroup() %>%
  select(occupation) %>%
  unique() %>%
  .[[1]]

data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    subcategories.order = descending_order_of_occupations,
                    value.column = ~total)
```

Frustratingly, at the time of writing `r Sys.Date()` the `plotly` library reverses the order of legends compared to bars:

```{r}
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "plotly",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    subcategories.order = descending_order_of_occupations,
                    value.column = ~total) %>%
    layout(margin = list(l = 150),
         yaxis = list(title = ""))
```

# Stacked Bar Charts

Stacked barcharts group data into categories and then create one bar with segments corresponding to the subcategory values. Note that there are two types of stacked:

- normal: the length of the bar for each category is the total of the length of each subcategory (i.e. within category total)
- percent: the length of all bars in the chart sum to 100% and the segment length corresponds to the within category percentage of values associated with that subcategory. See the [stacked percentage](#stacked-percentage) section for details.

```{r, echo=F}
descending_order_of_occupations <- data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  mutate(total = sum(count)) %>%
  mutate(total.in.group = sum(total)) %>%
  arrange(desc(total.in.group)) %>%
  ungroup() %>%
  select(occupation) %>%
  unique() %>%
  .[[1]]

descending_order_of_regions <- data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  mutate(total.in.group = sum(total)) %>%
  arrange(desc(total.in.group)) %>%
  select(country_group) %>%
  unique() %>%
  .[[1]]

data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    stacking.type = "normal",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    subcategories.order = descending_order_of_occupations,
                    value.column = ~total)
remove(descending_order_of_occupations, descending_order_of_regions)
```

The following will create a generic stacked bar chart with highcharter:

```{r}
library(highcharter)
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    stacking.type = "normal",
                    categories.column = ~country_group,
                    subcategories.column = ~occupation,
                    value.column = ~total)
```

It is convenient to enable grouped tooltips for `highcharter`, so that values for all subcategories are shown on hover. The following code will be used in all further examples with `highcharter` for stacked bar charts:

```{r}
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    stacking.type = "normal",
                    categories.column = ~country_group,
                    subcategories.column = ~occupation,
                    value.column = ~total) %>%
  hc_tooltip(shared = TRUE)
```


The following will create a generic stacked bar chart with plotly:

```{r}
library(plotly)
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "plotly",
                    stacking.type = "normal",
                    categories.column = ~country_group,
                    subcategories.column = ~occupation,
                    value.column = ~total)
```

At the time of writing `r Sys.Date()`, the plotly library does a poor job at ensuring the category labels (y-axis labels) are not chopped off. The `layout` function from `plotly` allows us to modify the margins and drop the unnecessary label for the y-axis.

```{r}
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "plotly",
                    stacking.type = "normal",
                    categories.column = ~country_group,
                    subcategories.column = ~occupation,
                    value.column = ~total) %>%
  layout(margin = list(l = 150),
         yaxis = list(title = ""))
```


## Ordering Categories

A good pattern to follow with stacked bar charts is to order the categories from "largest to smallest", i.e. the country_group with the most number of jobs at the top of the chart and the least at the bottom. This functionality is deliberately NOT provided by the `stacked_bar_chart` function, instead you *must* modify the data:

```{r}
descending_order_of_regions <- data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  mutate(total.in.group = sum(total)) %>%
  arrange(desc(total.in.group)) %>%
  select(country_group) %>%
  unique() %>%
  .[[1]]

data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "plotly",
                    stacking.type = "normal",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    value.column = ~total) %>%
  layout(margin = list(l = 150),
         yaxis = list(title = ""))
```

## Explicitly Ordering Subcategories

In addition to ordering the categories by descending size, it is useful to do the same for the subcategories:

```{r}
descending_order_of_occupations <- data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  mutate(total = sum(count)) %>%
  mutate(total.in.group = sum(total)) %>%
  arrange(desc(total.in.group)) %>%
  ungroup() %>%
  select(occupation) %>%
  unique() %>%
  .[[1]]

data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    stacking.type = "normal",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    subcategories.order = descending_order_of_occupations,
                    value.column = ~total) %>%
  hc_tooltip(shared = TRUE)
```


## Stacked Percentage {#stacked-percentage}

The `stacked_bar_chart` function automatically calculates within category percentages for subcategory values, for instance:

```{r}
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "highcharter",
                    stacking.type = "percent",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    subcategories.order = descending_order_of_occupations,
                    value.column = ~total) %>%
  hc_tooltip(shared = TRUE)
```

```{r}
data_stacked_bar_chart %>%
  group_by(country_group, occupation) %>%
  summarise(total = sum(count)) %>%
  ungroup() %>%
  stacked_bar_chart(library = "plotly",
                    stacking.type = "percent",
                    categories.column = ~country_group,
                    categories.order = descending_order_of_regions,
                    subcategories.column = ~occupation,
                    subcategories.order = descending_order_of_occupations,
                    value.column = ~total) %>%
  layout(margin = list(l = 150),
         yaxis = list(title = ""))
```


